### 2017-10-21 2-newfiles.Rmd

---
title: "New files"
author: "Florian Privé"
date: "20 octobre 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", out.width = "70%", 
                      fig.asp = 0.7, error = TRUE)
options(width = 85)
```

### Files for ttrees

```{r}
library(bigsnpr)
celiac2 <- snp_attach("backingfiles/celiacQC_sub1.rds")
G <- celiac2$genotypes
```

```{r}
# jdb
file.jdb <- "backingfiles/ttrees.jdb"
cat(rep(";", 5), sep = "\n", file = file.jdb)
big_apply(G, a.FUN = function(X, ind) {
  df <- cbind(paste0("ind_", ind), rep(3, length(ind)), as.data.frame(X[ind, ]))
  data.table::fwrite(df, file = file.jdb, sep = " ", append = TRUE,
                     quote = FALSE, col.names = FALSE)
  NULL
}, a.combine = "c", ind = rows_along(G), block.size = 1e3)
```

```{r}
# bloc
bigsnpr:::write.table2(
  bigstatsr:::CutBySize(ncol(G), block.size = 10)[, 1:2] - 1,
  file = file.bloc <- sub("\\.jdb$", ".bloc", file.jdb)
)
readLines(file.bloc, 5)
```

### File with recessive/dominant information

```{Rcpp}
// [[Rcpp::depends(bigstatsr, BH)]]
#include <bigstatsr/BMCodeAcc.h>

// [[Rcpp::export]]
void tripleBM(Environment BM, Environment BM2) {

  XPtr<FBM> xpMat = BM["address"];
  int n = xpMat->nrow();
  int m = xpMat->ncol();
  SubBMCode256Acc macc(xpMat, seq_len(n)-1, seq_len(m)-1, BM["code256"]);

  XPtr<FBM> xpMat2 = BM2["address"];
  BMAcc<unsigned char> macc2(xpMat2);

  int i, j, j2;

  for (j = j2 = 0; j < m; j++, j2 += 3) {
    for (i = 0; i < n; i++) {
      macc2(i, j2)   = macc(i, j);
      macc2(i, j2+1) = macc(i, j) >= 0.5;
      macc2(i, j2+2) = macc(i, j) >  1.5;
    }
  }
}
```

```{r}
snp_triple <- function(x) {

  G0 <- x$genotypes
  G <- G0$copy(code = round(G0$code256))

  newfile <- bigsnpr:::getNewFile(x, "tripled")
  G2 <- FBM.code256(nrow(G), 3 * ncol(G), 
                    code = G$code256,
                    backingfile = newfile,
                    save = TRUE)

  tripleBM(G, G2)

  paste0(newfile, ".rds")
}
```

```{r}
snp_triple(celiac2)
```

```{r}
# List all the files with their size in GB
round(sapply(list.files("backingfiles", full.names = TRUE), file.size) / 1024^3, 1)
```