### 2017-10-27 4-results.Rmd
save

---
title: "Results"
author: "Florian Privé"
date: "October 26, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", out.width = "70%", fig.asp = 0.7)
options(width = 110)
```

## Useful functions

```{r, message=FALSE}
library(tidyverse)
```

```{r}
myggplot <- function(..., coeff = 1) {
  bigstatsr:::MY_THEME(ggplot(...), coeff = coeff)
} 
```

```{r}
plot_results <- function(results, y, ylab = y) {

  dist <- "Distribution\nof effects"

  myggplot(results) +
    geom_boxplot(aes_string("method", y, color = "par.dist", 
                            fill = "par.dist"), alpha = 0.3) + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    facet_grid(par.model ~ par.causal) +
    theme(strip.text.x = element_text(size = rel(2)),
          strip.text.y = element_text(size = rel(2))) +
    labs(x = "Method", y = ylab, fill = dist, color = dist)
}
```

## Results with T-Trees

```{r}
top10 <- 1:110
top20 <- 1:220

# Put all results in a single tibble
results <- list.files("results1", full.names = TRUE) %>%
  map_dfr(~readRDS(.x)) %>%
  as_tibble() %>%
  mutate(
    par.causal = factor(map_chr(par.causal, ~paste(.x[1], .x[2], sep = " in ")),
                        levels = c("30 in HLA", paste(3 * 10^(1:3), "in all"))),
    AUC = map_dbl(eval, ~bigstatsr::AUC(.x[, 1], .x[, 2])),
    percCases10 = map_dbl(eval, ~mean(.x[order(.x[, 1], decreasing = TRUE)[top10], 2])),
    percCases20 = map_dbl(eval, ~mean(.x[order(.x[, 1], decreasing = TRUE)[top20], 2]))
  )
```

```{r}
results %>%
  filter(method %in% c("T-Trees", "logit-simple")) %>%
  group_by_at(c(vars(starts_with("par")), "method")) %>%
  summarise_at(c("timing", "nb.preds", "AUC", "percCases10", "percCases20"), mean) %>%
  print(n = Inf)
```

```{r, fig.width=16, out.width="95%", dev="svg"}
ttrees_vs_logit <- filter(results, method %in% c("T-Trees", "logit-simple"))

p_list <- list(
  plot_results(ttrees_vs_logit, "timing", "Timing (in seconds)") +
    scale_y_continuous(breaks = 0:10 * 2000, minor_breaks = NULL),
  plot_results(ttrees_vs_logit, "nb.preds", "Number of predictors (log-scale)") +
    scale_y_log10(breaks = c(10^(0:7), 3 * 10^(0:7)), minor_breaks = NULL,
                  labels = scales::comma_format()),
  plot_results(ttrees_vs_logit, "AUC") +
    scale_y_continuous(breaks = 0:10 / 10, minor_breaks = c(0:9 + 0.5) / 10),
  plot_results(ttrees_vs_logit, "percCases10", "Percentage of cases in top 10%") +
    scale_y_continuous(breaks = 0:10 / 10, minor_breaks = c(0:9 + 0.5) / 10)
)

lapply(p_list, function(p) p + theme(legend.position = "none")) %>%
  cowplot::plot_grid(plotlist = ., ncol = 2, align = "hv", scale = 0.9,
                     labels = LETTERS[1:4], label_size = 25) %>%
  cowplot::plot_grid(cowplot::get_legend(p_list[[1]]),
                     rel_widths = c(1, 0.15))
```

```{r}
ggsave("figures/ttrees.pdf", scale = 1/90, width = 1580, height = 1070)
```

## Results without T-Trees

### Correlation between predictive performance measures

```{r, fig.width=14, out.width="90%", dev="svg"}
cowplot::plot_grid(
  results %>%
    filter(par.h2 == 0.8) %>%
    myggplot(aes(AUC, percCases10, color = par.dist)) +
    geom_point() +
    geom_smooth(method = "lm") +
    theme(legend.position = c(0.8, 0.2)) +
    labs(y = "Percentage of cases in top 10%", 
         color = "Distribution\nof effects"),
  results %>%
    filter(par.h2 == 0.8) %>%
    myggplot(aes(AUC, percCases20, color = par.dist)) +
    geom_point() +
    geom_smooth(method = "lm") +
    theme(legend.position = c(0.8, 0.2)) +
    labs(y = "Percentage of cases in top 20%", 
         color = "Distribution\nof effects"),
  results %>%
    filter(par.h2 == 0.8) %>%
    myggplot(aes(AUC, percCases10, color = par.model)) +
    geom_point() +
    geom_smooth(method = "lm") +
    theme(legend.position = c(0.8, 0.2)) +
    scale_colour_brewer(palette = "Set1") +
    labs(y = "Percentage of cases in top 10%", 
         color = "Model"),
  results %>%
    filter(par.h2 == 0.8) %>%
    myggplot(aes(AUC, percCases20, color = par.model)) +
    geom_point() +
    geom_smooth(method = "lm") +
    theme(legend.position = c(0.8, 0.2)) +
    scale_colour_brewer(palette = "Set1") +
    labs(y = "Percentage of cases in top 20%", 
         color = "Model"),
  labels = LETTERS[1:4], align = "hv", label_size = 25, scale = 0.95
)
```

```{r}
ggsave("figures/AUC-corr.pdf", scale = 1/90, width = 1300, height = 950)
```