### 2017-11-04 1-preprocessing.Rmd
save

---
title: "Preprocessing"
author: "Florian Privé"
date: "October 20, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", out.width = "70%", 
                      fig.asp = 0.7, error = TRUE)
options(width = 85)
```

### QC and imputation

This was done in [another paper](https://doi.org/10.1101/190926).

```{r}
file.copy(paste0("../paper1-packages/backingfiles/celiacQC", c(".bk", ".rds")),
          paste0("backingfiles/celiacQC", c(".bk", ".rds")))

library(bigsnpr)
str(celiac <- snp_attach("backingfiles/celiacQC.rds"), max.level = 1)
```

### Add population from external files

```{r}
# Get population
pop.files <- list.files(path = "../thesis-celiac/Dubois2010_data/", 
                        pattern = "cluster_*", full.names = TRUE)
pop <- snp_getSampleInfos(celiac, pop.files)[[1]]
pop.names <- c("Netherlands", "Italy", "UK1", "UK2", "Finland")
celiac$fam$family.ID <- pop.names[pop]
celiac <- snp_save(celiac)
# Read modified file to check
celiac <- snp_attach("backingfiles/celiacQC.rds")
rle(celiac$fam$family.ID)
```

### Filter samples based on population structure

```{r}
# Compute PCA
obj.svd <- snp_autoSVD(celiac$genotypes, 
                       celiac$map$chromosome,
                       celiac$map$physical.pos, 
                       ncores = nb_cores())

library(ggplot2)
plot(obj.svd, type = "scores", scores = 1:2) + 
  aes(color = celiac$fam$family.ID) + 
  labs(color = "Population")
```

```{r}
# Which are UK controls?
ind.UKcontrols <- which(startsWith(celiac$fam$family.ID, "UK") & 
                          celiac$fam$affection == 1)
obj.svd$u[-ind.UKcontrols, ] <- NA
```

```{r, warning=FALSE}
# Compute Mahalanobis distance on PCs
distM <- rep(NA, nrow(celiac$genotypes))
distM[ind.UKcontrols] <- robust::covRob(obj.svd$u[ind.UKcontrols, ])$dist
THR <- 30
# Plot UK controls, coloring with Mahalanobis distance
library(magrittr)
p_list <- lapply(list(1:2, 3:4, 5:6, 7:8), function(scores) {
  plot(obj.svd, type = "scores", scores = scores, coeff = 0.7) + 
    aes(color = (distM > THR)) + 
    labs(color = "Outlier?") +
    scale_color_manual(breaks = c(TRUE, FALSE), values = scales::hue_pal()(2))
}) 
lapply(p_list, function(p) p + theme(legend.position = "none")) %>%
  cowplot::plot_grid(plotlist = ., ncol = 2, align = "hv", scale = 0.95) %>%
  cowplot::plot_grid(cowplot::get_legend(p_list[[1]]),
                     rel_widths = c(1, 0.15))
```

```{r}
ggsave("figures/outliers-pop.pdf", scale = 1/90, width = 870, height = 650)
```

```{r}
# New dataset with only non-outliers UK controls
length(ind.keep <- which(distM < THR))
celiac2 <- snp_attach(subset(celiac, ind.row = ind.keep))
obj.svd2 <- snp_autoSVD(celiac2$genotypes, 
                        celiac2$map$chromosome,
                        celiac2$map$physical.pos,
                        ncores = nb_cores())
plot(obj.svd2)
plot(obj.svd2, type = "scores") 
# Save this new PCA
saveRDS(obj.svd2, "backingfiles/PCA2.rds")
```

```{r}
ggsave("figures/no-pop.pdf", scale = 1/80, width = 620, height = 570)
```