### 2017-12-13 3-oracle.Rmd
save

---
title: "Oracle"
author: "Florian Privé"
date: "November 17, 2017"
output: 
  html_document:
    code_folding: show
bibliography: paper/refs.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", out.width = "70%", 
                      fig.asp = 0.7, dev = "svg")
options(width = 110)
```

## Data and useful functions

```{r, message=FALSE}
library(tidyverse)
library(bigsnpr)
library(foreach)
celiac2 <- snp_attach("backingfiles/celiacQC_sub1.rds")
CHR <- celiac2$map$chromosome
POS <- celiac2$map$physical.pos
(G <- celiac2$genotypes)
ind.HLA <- snp_indLRLDR(CHR, POS, subset(LD.wiki34, ID == "hild12"))
NCORES <- nb_cores()
```

```{r}
myggplot <- function(..., coeff = 1) {
  bigstatsr:::MY_THEME(ggplot(...), coeff = coeff)
} 
```

```{r}
boot <- function(x, n = 1e5, f = mean) {
  sd(replicate(n, f(sample(x, replace = TRUE))))
}
```

## Get maximum AUCs

```{r}
K <- 0.3
```

### Estimation from Wray 2010

Computation of the maximum AUC from equation (3) of @wray2010genetic. Approximation is less accurate for high heritabilities.

```{r}
sapply(c(0.5, 0.8), function(h2) {
  T0 <- qnorm(1 - K)
  z <- dnorm(T0)
  i <- z / K
  v <- -i * K / (1 - K)
  num <- (i - v) * h2
  deno.part <- 1 - h2 * i * (i - T0) + (1 - h2 * v * (v - T0))
  pnorm(num / sqrt(h2 * deno.part))
})
```

### Common estimation for all simulations

Assuming that $y \sim N(0, h^2)$ and knowing that $\epsilon \sim N(0, 1-h^2)$, we can estimate the theoretical value of the AUC that can be achieved given the heritability $h^2$ and a liability threshold model on $y + \epsilon$ with a prevalence of 30%.

```{r}
nsimu <- 1e7

doParallel::registerDoParallel(NCORES)
aucs <- foreach(i = seq_len(100), .combine = "cbind") %:%
  foreach(h2 = c(0.5, 0.8), .combine = "rbind") %dopar% {
    S <- rnorm(nsimu, mean = 0, sd = sqrt(h2))
    E <- rnorm(nsimu, mean = 0, sd = sqrt(1 - h2))
    Y <- (S + E) > qnorm(1 - K)

    bigstatsr::AUC(S, Y)
  }
doParallel::stopImplicitCluster()
apply(aucs, 1, summary)
```

### Different estimations for all simulations

```{r}
get_oracle <- function(
  G,                                        ## matrix of genotypes
  h2,                                       ## heritability 
  M,                                        ## nbs of causal variants
  ind.possible = cols_along(G),             ## indices of possible causal variants
  effects.dist = c("gaussian", "laplace"),  ## distribution of effects 
  model = c("simple", "fancy"),             ## model for simulation
  K = 0.3                                   ## prevalence
) {

  effects.dist  <- match.arg(effects.dist)
  model <- match.arg(model)

  set <- sample(ind.possible, size = M)
  effects <- `if`(effects.dist == "gaussian", 
                  rnorm(M, sd = sqrt(h2 / M)),
                  rmutil::rlaplace(M, s = sqrt(h2 / (2*M))))

  if (model == "simple") {
    # only linear
    y.simu <- scale(G[, set]) %*% effects
  } else {
    sets <- split(1:M, sample(rep_len(1:3, M)))
    # linear
    ind1 <- sets[[1]]
    y.simu <- scale(G[, set[ind1]]) %*% effects[ind1]
    # recessive / dominant
    ind2 <- sets[[2]]
    y.simu <- y.simu + scale(G[, set[ind2]] > 0.5) %*% effects[ind2]
    # interactions
    ind3 <- matrix(sets[[3]], ncol = 2)
    y.simu <- y.simu + scale(G[, set[ind3[, 1]]] * G[, set[ind3[, 2]]]) %*% 
      effects[ind3[, 1]] * sqrt(2)
  }

  y.simu <- y.simu / sd(y.simu) * sqrt(h2)
  stopifnot(all.equal(drop(var(y.simu)), h2))
  S <- y.simu
  y.simu <- S + rnorm(nrow(G), sd = sqrt(1 - h2))
  pheno <- as.numeric(y.simu > qnorm(1 - K))
  AUC(S, pheno)
}
```

```{r}
params.grid2 <- expand.grid(
  par.causal = list(c(30, "all"), c(300, "all"), c(3000, "all"), c(30, "HLA")), 
  par.dist   = c("gaussian", "laplace"), 
  par.h2     = c(0.5, 0.8), 
  par.model  = c("simple", "fancy"),
  num.simu   = 1:100,
  stringsAsFactors = FALSE
) 
```

```{r}
AUC_oracle <- function(G, params, ind.HLA) {

  par.causal <- params[["par.causal"]][[1]]

  # Simulate phenotypes
  AUC <- get_oracle(
    G,    
    h2 = params[["par.h2"]], 
    M = as.integer(par.causal[1]), 
    ind.possible = `if`(par.causal[2] == "all", cols_along(G), ind.HLA),
    effects.dist = params[["par.dist"]], 
    model = params[["par.model"]], 
    K = 0.3                 
  )

  params$AUC <- AUC
  params
}
```

#### On all 22 chromosomes

```{r}
doParallel::registerDoParallel(NCORES)
oracles <- foreach(i = rows_along(params.grid2), .combine = "rbind") %dopar% {
  AUC_oracle(G, params.grid2[i, ], ind.HLA)
}
doParallel::stopImplicitCluster()
```

```{r}
oracles %>% 
  mutate(
    par.causal = factor(map_chr(par.causal, ~paste(.x[1], .x[2], sep = " in ")),
                        levels = c("30 in HLA", paste(3 * 10^(1:3), "in all")))) %>%
  group_by(par.causal, par.dist, par.h2, par.model) %>%
  summarise(AUC_mean = round(mean(AUC), 4),
            AUC_mean_boot = round(boot(AUC), 4)) %>%
  print(n = Inf)
```

#### On only the chromosome 6

```{r}
G6 <- big_copy(G, ind.col = which(CHR == 6))
ind.HLA6 <- snp_indLRLDR(CHR[CHR == 6], POS[CHR == 6], 
                         subset(LD.wiki34, ID == "hild12"))

doParallel::registerDoParallel(NCORES)
oracles_chr6 <- foreach(i = rows_along(params.grid2), .combine = "rbind") %dopar% {
  AUC_oracle(G6, params.grid2[i, ], ind.HLA6)
}
doParallel::stopImplicitCluster()
```

```{r}
tmp <- oracles_chr6 %>% 
  mutate(
    par.causal = factor(map_chr(par.causal, ~paste(.x[1], .x[2], sep = " in ")),
                        levels = c("30 in HLA", paste(3 * 10^(1:3), "in all")))) %>%
  group_by(par.causal, par.dist, par.h2, par.model) %>%
  summarise(AUC_mean = round(mean(AUC), 4),
            AUC_mean_boot = round(boot(AUC), 4)) %>%
  print(n = Inf)
```

## References