### 2018-03-01 stability-CMSA.Rmd
add celiac results

## Stability of selection of the "logit-simple"

Train on 100 bootstrap samples of the Celiac dataset.

```{r, eval=FALSE}
if (!dir.exists("results6")) dir.create("results6")

for (i in seq_len(100)) {

  res.file <- paste0("results6/simu_", i, ".rds")
  if (file.exists(res.file)) next

  ind.train <- sort(sample(nrow(G), replace = TRUE))

  cmsa.logit <- big_CMSA(FUN = big_spLogReg, feval = AUC,
                         X = G, y.train = y[ind.train], 
                         ind.train = ind.train, 
                         covar.train = obj.svd$u[ind.train, ],
                         alpha = 0.5, dfmax = 20e3, 
                         ncores = NCORES)

  saveRDS(list(betas = cmsa.logit[cols_along(G)],
               ind.test = setdiff(rows_along(G), ind.train)),
          file = res.file)
}
```

```{r}
results6 <- list.files("results6", full.names = TRUE) %>%
  map(~readRDS(.x)) %>%
  do.call("cbind", .)
```

```{r}
n_enter <- rowSums(results6 != 0)
mean(n_enter == 0)                ## Proportion of SNPs never entering
n_enter_once <- n_enter[n_enter != 0]
mean(n_enter_once)                ## Mean number of times a SNPs enter 
                                  ## if entering at least once
bigstatsr:::MY_THEME(
  qplot(n_enter_once, geom = "bar") +
    labs(x = "Number of times")
)
```

```{r}
ggsave("figures/supp-celiac-stabilitybar.pdf", scale = 1/90, width = 1000, height = 700)
```

```{r}
length(always_enter <- which(n_enter == ncol(results6)))
cowplot::plot_grid(
  snp_manhattan(gwas.train.gc, CHR, POS, labels = labels, 
                ind.highlight = always_enter),
  snp_manhattan(gwas.train.gc, CHR, POS, labels = labels, 
                ind.highlight = always_enter) +
    coord_cartesian(ylim = c(0, 25)) + 
    geom_hline(yintercept = -log10(5e-8), color = "red", linetype = 3),
  align = "hv", ncol = 1, labels = LETTERS[1:2], label_size = 25, scale = 0.95
)
```

```{r}
ggsave("figures/supp-celiac-man-col.png", scale = 1/90, width = 1300, height = 1000)
```

```{r}
sq_effects_always <- colSums((results6[always_enter, ])^2)
sq_effects_others <- colSums((results6[-always_enter, ])^2)
summary(100 * sq_effects_always / (sq_effects_always + sq_effects_others))
```