### 2018-03-02 5-results-celiac.Rmd
need to add multiple results

---
title: "Analysis of celiac disease"
author: "Florian Privé"
date: "March 1, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", out.width = "70%", fig.width = 10,
                      fig.asp = 0.7, dev = "svg", cache = TRUE)
options(width = 95)
```

```{r}
library(tidyverse)
myggplot <- function(..., coeff = 1) {
  bigstatsr:::MY_THEME(ggplot(...), coeff = coeff)
} 
```

## Data

```{r, cache=FALSE}
# Read file, see "Preprocessing" notebook
library(bigsnpr)
celiac <- snp_attach("backingfiles/celiacQC.rds")
G <- celiac$genotypes
CHR <- celiac$map$chromosome
POS <- celiac$map$physical.pos
NCORES <- nb_cores()
y <- celiac$fam$affection - 1
```

```{r}
# Compute PCA
obj.svd <- snp_autoSVD(G, CHR, POS, ncores = NCORES)
```

## Methods' functions

Each method's function returns a tibble (data frame) with 4 columns:

1. the name of the method,
2. the predictive scores and true phenotypes for the test set, as a [list-column](http://r4ds.had.co.nz/many-models.html),
3. the timing of the main computations (in seconds),
4. the number of SNPs used for the prediction.

### PRS

```{r}
PRS <- function(G, infos.chr, infos.pos, 
                pheno.all, covar.all, 
                ind.train, ind.test) {

  timing <- system.time({

    # GWAS
    gwas.train <- big_univLogReg(
      G, pheno.all[ind.train], ind.train = ind.train, 
      covar.train = covar.all[ind.train, , drop = FALSE], 
      ncores = nb_cores()
    )
    gwas.train.gc <- snp_gc(gwas.train)

    # Clumping on the test set
    ind.keep <- snp_clumping(G, infos.chr = infos.chr,
                             ind.row = ind.test,
                             thr.r2 = 0.2, 
                             S = abs(gwas.train.gc$score),
                             size = 500,
                             is.size.in.bp = TRUE,
                             infos.pos = infos.pos,
                             ncores = nb_cores())

    # PRS
    thrs <- c(0, -log10(5e-8), exp(seq(log(0.1), log(100), length.out = 100)))
    lpS <- -predict(gwas.train.gc)
    nb.pred <- sapply(thrs, function(thr) sum(lpS[ind.keep] > thr))
    prs <- snp_PRS(G, betas.keep = gwas.train.gc$estim[ind.keep],
                   ind.test = ind.test,
                   ind.keep = ind.keep,
                   lpS.keep = lpS[ind.keep], 
                   thr.list = thrs)
  })[3]

  ind.best <- which.max(apply(prs, 2, AUC, target = pheno.all[ind.test]))

  tibble(
    method = "PRS-max",
    eval = list(cbind(prs[, ind.best], pheno.all[ind.test])),
    timing = timing,
    nb.preds = nb.pred[ind.best]
  )
}
```

### Logistic regression

```{r}
logit.CMSA <- function(G, pheno.all, covar.all, ind.train, ind.test, method, print_all = FALSE) {

  timing <- system.time({

    cmsa.logit <- big_spLogReg(X = G, y01.train = pheno.all[ind.train], 
                               ind.train = ind.train, 
                               covar.train = covar.all[ind.train, , drop = FALSE],
                               ncores = nb_cores())

    preds.all <- predict(cmsa.logit, X = G, ind.row = ind.test, 
                         covar.row = covar.all[ind.test, , drop = FALSE])
    if (print_all) print(preds.all)
    preds <- rowMeans(preds.all)
  })[3]

  tibble(
    method   = method, 
    eval     = list(cbind(preds, pheno.all[ind.test])),
    timing   = timing,
    nb.preds = sum(rowSums(sapply(cmsa.logit, function(x) x$beta.X) != 0) > 0)
  )
}
```

## Predictive models and plots

```{r}
# Divide in training/test sets
dim(G)
set.seed(1)
ind.train <- sort(sample(nrow(G), size = 12e3))
ind.test <- setdiff(rows_along(G), ind.train)
```

### PRS

```{r}
res.PRS <- PRS(
  G = G,
  infos.chr = CHR,
  infos.pos = POS,
  pheno.all = y,
  covar.all = obj.svd$u,
  ind.train = ind.train,
  ind.test = ind.test
)
# GWAS
system.time(
  gwas.train <- big_univLogReg(
    G, y[ind.train], ind.train = ind.train, 
    covar.train = obj.svd$u[ind.train, , drop = FALSE], 
    ncores = NCORES
  )
)
```

```{r, dev='png', fig.asp=0.8}
# Q-Q plot
snp_qq(gwas.train) +
  coord_cartesian(ylim = c(0, 15))
gwas.train.gc <- snp_gc(gwas.train)
```

```{r, dev="png", fig.asp=0.9, out.width="90%"}
# Manhattan plot
labels <- 1:22; labels[c(11, 13, 15, 17, 18, 20, 21)] <- ""
cowplot::plot_grid(
  snp_manhattan(gwas.train.gc, CHR, POS, labels = labels),
  snp_manhattan(gwas.train.gc, CHR, POS, labels = labels) +
    coord_cartesian(ylim = c(0, 22)) + 
    geom_hline(yintercept = -log10(5e-8), color = "red", linetype = 3),
  align = "hv", ncol = 1, labels = LETTERS[1:2], label_size = 25, scale = 0.95
)
```

```{r}
# Clumping on the test set
ind.keep <- snp_clumping(G, infos.chr = CHR,
                         ind.row = ind.test,
                         thr.r2 = 0.2, 
                         S = abs(gwas.train.gc$score),
                         size = 500,
                         is.size.in.bp = TRUE,
                         infos.pos = POS,
                         ncores = NCORES)
```

```{r}
# PRS
thrs <- c(0, -log10(5e-8), exp(seq(log(0.1), log(100), length.out = 100)))
lpS <- -predict(gwas.train.gc)
prs <- snp_PRS(G, betas.keep = gwas.train.gc$estim[ind.keep],
               ind.test = ind.test,
               ind.keep = ind.keep,
               lpS.keep = lpS[ind.keep], 
               thr.list = thrs)
nb.pred <- sapply(thrs, function(thr) sum(lpS[ind.keep] > thr))
aucs <- apply(prs, 2, AUC, target = y[ind.test])
summary(aucs)
nb.pred[which.max(aucs)]
```

```{r}
pAUC <- function(pred, target, p = 0.1) {
  val.min <- min(target)
  q <- quantile(pred[target == val.min], probs = 1 - p)
  ind <- (target != val.min) | (pred > q)
  tryCatch(bigstatsr::AUC(pred[ind], target[ind]) * p, 
           error = function(e) p^2 / 2)  # all same prediction
}
plot_density_AUC <- function(pred, true = y[ind.test]) {
  data.frame(pred = pred, pheno = as.factor(true)) %>%
  myggplot() +
  geom_density(aes(pred, fill = pheno), alpha = 0.4) +
  ggtitle(sprintf("AUC = %s%% | pAUC = %s", 
                  round(100 * AUC(pred, true), 2),
                  round(pAUC(pred, true), 4))) + 
  theme(plot.title = element_text(size = rel(1.5), hjust = 0.5))
}
(p1 <- plot_density_AUC(prs[, which.max(aucs)]))
```

### Logistic regression

Cross-Model Selection and Averaging (CMSA):

```{r}
# Train logit-simple
system.time(
  logit <- big_spLogReg(X = G, y01.train = y[ind.train], 
                        ind.train = ind.train, 
                        covar.train = obj.svd$u[ind.train, ],
                        ncores = NCORES)
)
# Get K = 10 predictions/individuals
preds1 <- predict(logit, X = G, ind.row = ind.test, 
                  covar.row = obj.svd$u[ind.test, ])
(nb.pred1 <- sum(rowSums(sapply(logit, function(l) l$beta.X != 0)) > 0))
(aucs1 <- apply(preds1, 2, AUC, target = y[ind.test]))
# Combining all K predictions
preds1.comb <- rowMeans(preds1)
AUC(preds1.comb, y[ind.test])
```

```{r}
(p2 <- plot_density_AUC(preds1.comb))
```

### Logistic regression with feature engineering

```{r}
G2 <- big_attach("backingfiles/celiacQC_tripled1.rds")
system.time(
  logit2 <- big_spLogReg(X = G2, y01.train = y[ind.train], 
                         ind.train = ind.train, 
                         covar.train = obj.svd$u[ind.train, ],
                         ncores = NCORES)
)

preds2 <- predict(logit2, X = G2, ind.row = ind.test, 
                  covar.row = obj.svd$u[ind.test, ])
(nb.pred2 <- sum(rowSums(sapply(logit2, function(l) l$beta.X != 0)) > 0))
(aucs2 <- apply(preds2, 2, AUC, target = y[ind.test]))
# Combining all K predictions
preds2.comb <- rowMeans(preds2)
AUC(preds2.comb, y[ind.test])
```

```{r}
(p3 <- plot_density_AUC(preds2.comb))
```

```{r, out.width="70%", fig.asp=900/780}
list(p1, p2, p3) %>%
  lapply(function(p) p + theme(legend.position = "none")) %>%
  cowplot::plot_grid(plotlist = ., ncol = 1, labels = LETTERS[1:3], label_size = 20, scale = 0.95) %>%
  cowplot::plot_grid(., cowplot::get_legend(p1), rel_widths = c(1, 0.15), scale = 0.95)
ggsave("figures/supp-score-densities.pdf", scale = 1/100, width = 780, height = 900)
```

### ROC Curves

```{r, fig.asp=1, fig.width=8}
library(plotROC)
scores_tidy <- tibble(
  d = y[ind.test], 
  "PRS-max"      = prs[, which.max(aucs)],
  "logit-simple" = preds1.comb,
  "logit-triple" = preds2.comb
) %>%
  gather(key = "Method", value = "Score", -d)

bigstatsr:::MY_THEME(
  ggplot(scores_tidy, aes(d = d, m = Score, color = Method, linetype = Method)) +
    style_roc(xlab = "1 - Specificity", ylab = "Sensitivity")
) +
  geom_roc(n.cuts = 0, size = 2) +
  theme(legend.position = c(0.7, 0.3), legend.key.width = unit(4, "line")) +
  coord_equal()
```

```{r}
ggsave("figures/celiac-roc.pdf", scale = 1/90, width = 700, height = 700)
```

## Predictive methods (multiple times)

```{r}
# Divide in training/test sets
ind.train <- sort(sample(nrow(G), size = 12e3))
ind.test <- setdiff(rows_along(G), ind.train)
```

### PRS

```{r}
# PRS
PRS <- function() {

  system.time(
  gwas.train <- big_univLogReg(
    G, y[ind.train], ind.train = ind.train, 
    covar.train = obj.svd$u[ind.train, , drop = FALSE], 
    ncores = NCORES
  )
)
}

```

### Logistic regression

Cross-Model Selection and Averaging (CMSA):

```{r}
# Train logit-simple
system.time(
  logit <- big_spLogReg(X = G, y01.train = y[ind.train], 
                        ind.train = ind.train, 
                        covar.train = obj.svd$u[ind.train, ],
                        ncores = NCORES)
)
# Get K = 10 predictions/individuals
preds1 <- predict(logit, X = G, ind.row = ind.test, 
                  covar.row = obj.svd$u[ind.test, ])
(nb.pred1 <- sum(rowSums(sapply(logit, function(l) l$beta.X != 0)) > 0))
(aucs1 <- apply(preds1, 2, AUC, target = y[ind.test]))
# Combining all K predictions
preds1.comb <- rowMeans(preds1)
AUC(preds1.comb, y[ind.test])
```

```{r}
(p2 <- plot_density_AUC(preds1.comb))
```

### Logistic regression with feature engineering

```{r}
G2 <- big_attach("backingfiles/celiacQC_tripled1.rds")
system.time(
  logit2 <- big_spLogReg(X = G2, y01.train = y[ind.train], 
                         ind.train = ind.train, 
                         covar.train = obj.svd$u[ind.train, ],
                         ncores = NCORES)
)

preds2 <- predict(logit2, X = G2, ind.row = ind.test, 
                  covar.row = obj.svd$u[ind.test, ])
(nb.pred2 <- sum(rowSums(sapply(logit2, function(l) l$beta.X != 0)) > 0))
(aucs2 <- apply(preds2, 2, AUC, target = y[ind.test]))
# Combining all K predictions
preds2.comb <- rowMeans(preds2)
AUC(preds2.comb, y[ind.test])
```

```{r}
(p3 <- plot_density_AUC(preds2.comb))
```

```{r, out.width="70%", fig.asp=900/780}
list(p1, p2, p3) %>%
  lapply(function(p) p + theme(legend.position = "none")) %>%
  cowplot::plot_grid(plotlist = ., ncol = 1, labels = LETTERS[1:3], label_size = 20, scale = 0.95) %>%
  cowplot::plot_grid(., cowplot::get_legend(p1), rel_widths = c(1, 0.15), scale = 0.95)
ggsave("figures/supp-score-densities.pdf", scale = 1/100, width = 780, height = 900)
```

## Top scores

First, we get the most significant SNP.

```{r}
# Most significant SNP
theone <- which.max(-predict(gwas.train.gc))
ind0 <- which(G[ind.test, theone] == 0)
(M <- length(ind0))
table(Geno = G[, theone], forcats::fct_recode(as.factor(y), Control = "0", Case = "1"))
```

```{r}
# one geno
mean(y[ind.test[ind0]]) 
# prs max
indPRS <- order(prs[, which.max(aucs)], decreasing = TRUE)[1:M]
mean(y[ind.test[indPRS]])
# logit-simple (CMSA)
indLS <- order(preds1.comb, decreasing = TRUE)[1:M]
mean(y[ind.test[indLS]])
# logit-triple (CMSA)
indLT <- order(preds2.comb, decreasing = TRUE)[1:M]
mean(y[ind.test[indLT]])
```