### 2018-04-26 3-simus.Rmd
add biglasso comparison

---
title: "Simulations"
author: "Florian Privé"
date: "October 20, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", out.width = "70%", 
                      fig.asp = 0.7)
options(width = 85)
```

## Methods' functions

```{r, message=FALSE}
library(tidyverse)
library(glue)
NCORES <- 6
```

Each method's function returns a tibble (data frame) with 4 columns:

1. the name of the method,
2. the predictive scores and true phenotypes for the test set, as a [list-column](http://r4ds.had.co.nz/many-models.html),
3. the timing of the main computations (in seconds),
4. the number of SNPs used for the prediction.

### PRS

```{r}
10^(-c(0, -log10(5e-8), exp(seq(log(0.1), log(100), length.out = 100)))) %>%
  sprintf("%.1e", .) %>%
  paste(collapse = ", ")
```

```{r}
PRS <- function(G, infos.chr, infos.pos, 
                pheno.all, covar.all, 
                ind.train, ind.test) {

  timing <- system.time({

    # GWAS
    gwas.train <- big_univLogReg(
      G, pheno.all[ind.train], ind.train = ind.train, 
      covar.train = covar.all[ind.train, , drop = FALSE], 
      ncores = NCORES
    )
    gwas.train.gc <- snp_gc(gwas.train)

    # Clumping on the test set
    ind.keep <- snp_clumping(G, infos.chr = infos.chr,
                             ind.row = ind.test,
                             thr.r2 = 0.2, 
                             S = abs(gwas.train.gc$score),
                             size = 500,
                             is.size.in.bp = TRUE,
                             infos.pos = infos.pos,
                             ncores = NCORES)

    # PRS
    thrs <- c(0, -log10(5e-8), exp(seq(log(0.1), log(100), length.out = 100)))
    lpS <- -predict(gwas.train.gc)
    nb.pred <- sapply(thrs, function(thr) sum(lpS[ind.keep] > thr))
    prs <- snp_PRS(G, betas.keep = gwas.train.gc$estim[ind.keep],
                   ind.test = ind.test,
                   ind.keep = ind.keep,
                   lpS.keep = lpS[ind.keep], 
                   thr.list = thrs)
  })[3]

  ind.best <- which.max(apply(prs, 2, AUC, target = pheno.all[ind.test]))

  methods <- c("PRS-all", "PRS-stringent", "PRS-max")
  indices <- c(1:2, ind.best)

  lapply(1:3, function(i) {
    k <- indices[i]
    tibble(
      method = methods[i],
      eval = list(cbind(prs[, k], pheno.all[ind.test])),
      timing = timing,
      nb.preds = nb.pred[k]
    )
  }) %>%
    bind_rows()
}
```

### Logistic regression

```{r}
logit.CMSA <- function(G, pheno.all, covar.all, ind.train, ind.test, method) {

  timing <- system.time({

    cmsa.logit <- big_spLogReg(X = G, y01.train = pheno.all[ind.train], 
                               ind.train = ind.train, 
                               covar.train = covar.all[ind.train, , drop = FALSE],
                               ncores = NCORES)

    preds <- rowMeans(
      predict(cmsa.logit, X = G, ind.row = ind.test, 
                     covar.row = covar.all[ind.test, , drop = FALSE])
    )
  })[3]

  tibble(
    method   = method, 
    eval     = list(cbind(preds, pheno.all[ind.test])),
    timing   = timing,
    nb.preds = sum(rowSums(sapply(cmsa.logit, function(x) x$beta.X) != 0) > 0)
  )
}
```

### T-Trees

```{r}
TTree <- "../../TTree-source/TTree"

ttrees <- function(TTree, file.base, pheno.all, ind.train, ind.test,
                   n.trees = 100) {

  file0.jdb  <- paste0(file.base, ".jdb")
  file0.bloc <- paste0(file.base, ".bloc")

  # Write jdb file with new pheno
  tmpfile <- tempfile()
  write(c(rep("", 5), pheno.all), ncolumns = 1, tmpfile)
  # https://stackoverflow.com/a/7846550/6103040
  system(sprintf("awk 'FNR==NR{a[NR]=$1;next}{$2=a[FNR]}1' %s %s > %s",
                 tmpfile, file0.jdb, file.jdb <- paste0(tmpfile, ".jdb")))
  # Write indices of learning and validation sets
  file.learn <- paste0(tmpfile, "_learn.txt")
  file.val   <- paste0(tmpfile, "_val.txt")
  cat(ind.train - 1, file = file.learn, sep = "\t")
  cat(ind.test - 1,  file = file.val,   sep = "\t")

  timing <- system.time(
    system(glue::glue(
      "{TTree}",
      " -j {file.jdb}",
      " -m 3",
      " -b {file0.bloc}",
      " -l {file.learn}",
      " -v {file.val}", 
      " -t {n.trees} -k 1000 -c 5 -n 2000",
      " -x -s"
    ))
  )[3]

  file.roc <- sprintf("%s_k1000_m3_t%d_ic5_nmin2000_0000.roc",
                      file.jdb, n.trees)
  file.vim <- sub("\\.roc$", ".vim", file.roc)
  preds <- read.table(file.roc, header = FALSE)

  tibble(
    method   = "T-Trees",
    eval     = list(preds[match(ind.test - 1, preds[[1]]), 2:3]),
    timing   = timing,
    nb.preds = sum(read.table(file.vim)$V2 != 0)
  )
}
```

### Logistic regression with biglasso

```{r}
logit.biglasso <- function(G3, pheno.all, covar.all, ind.train, ind.test) {

  timing <- system.time({

    biglasso <- biglasso(G3, pheno.all, ind.train, ncores = NCORES,
                         penalty = "enet", family = "binomial", alpha = 0.5)

    preds <- 1 / (1 + exp(-predict(biglasso, G3, ind.test)))
  })[3]

  aucs <- apply(preds, 2, bigstatsr::AUC, target = pheno.all[ind.test])
  ind.max <- which.max(aucs)

  tibble(
    method   = "biglasso", 
    eval     = list(cbind(preds[, ind.max], pheno.all[ind.test])),
    timing   = timing,
    nb.preds = sum(biglasso$beta[, ind.max] != 0)
  )
}
```

## Simulations

### Data

```{r}
library(bigsnpr)
celiac2 <- snp_attach("backingfiles/celiacQC_sub1.rds")
CHR <- celiac2$map$chromosome
POS <- celiac2$map$physical.pos
(G <- celiac2$genotypes)
(G2 <- big_attach("backingfiles/celiacQC_sub1_tripled1.rds"))

covar.all <- readRDS("backingfiles/PCA2.rds")$u

n <- nrow(G)
ind.HLA <- snp_indLRLDR(CHR, POS, subset(LD.wiki34, ID == "hild12"))
```

### Simulation of phenotypes

```{r}
get_pheno <- function(
  G,                                        ## matrix of genotypes
  h2,                                       ## heritability 
  M,                                        ## nbs of causal variants
  ind.possible = cols_along(G),             ## indices of possible causal variants
  effects.dist = c("gaussian", "laplace"),  ## distribution of effects 
  model = c("simple", "fancy"),             ## model for simulation
  K = 0.3                                   ## prevalence
) {

  effects.dist  <- match.arg(effects.dist)
  model <- match.arg(model)

  set <- sample(ind.possible, size = M)
  effects <- `if`(effects.dist == "gaussian", 
                  rnorm(M, sd = sqrt(h2 / M)),
                  rmutil::rlaplace(M, s = sqrt(h2 / (2*M))))

  if (model == "simple") {
    # only linear
    y.simu <- scale(G[, set]) %*% effects
  } else {
    sets <- split(1:M, sample(rep_len(1:3, M)))
    # linear
    ind1 <- sets[[1]]
    y.simu <- scale(G[, set[ind1]]) %*% effects[ind1]
    # recessive / dominant
    ind2 <- sets[[2]]
    y.simu <- y.simu + scale(G[, set[ind2]] > 0.5) %*% effects[ind2]
    # interactions
    ind3 <- matrix(sets[[3]], ncol = 2)
    y.simu <- y.simu + scale(G[, set[ind3[, 1]]] * G[, set[ind3[, 2]]]) %*% 
      effects[ind3[, 1]] * sqrt(2)
  }

  y.simu <- y.simu / sd(y.simu) * sqrt(h2)
  stopifnot(all.equal(drop(var(y.simu)), h2))
  y.simu <- y.simu + rnorm(nrow(G), sd = sqrt(1 - h2))
  pheno <- as.numeric(y.simu > qnorm(1 - K))
}
```

### Scenario n°1 (with T-Trees)

```{r}
params.grid1 <- expand.grid(
  n.train    = 6000,
  par.causal = list(c(30, "all"), c(300, "all"), c(3000, "all"), c(30, "HLA")), 
  par.dist   = c("gaussian", "laplace"), 
  par.h2     = 0.8, 
  par.model  = c("simple", "fancy"),
  num.simu   = 1:5,
  stringsAsFactors = FALSE
) 
```

```{r, eval=TRUE}
if (!dir.exists("results1")) dir.create("results1")

for (i in rows_along(params.grid1)) {

  res.file <- paste0("results1/simu_", i, ".rds")
  if (file.exists(res.file)) next

  params <- params.grid1[i, ]
  par.causal <- params[["par.causal"]][[1]]

  # Simulate phenotypes
  pheno.all <- get_pheno(
    G,    
    h2 = params[["par.h2"]], 
    M = as.integer(par.causal[1]), 
    ind.possible = `if`(par.causal[2] == "all", cols_along(G), ind.HLA),
    effects.dist = params[["par.dist"]], 
    model = params[["par.model"]], 
    K = 0.3                 
  )

  # Split in training/test sets
  ind.train <- sort(sample(n, size = params[["n.train"]]))
  ind.test <- setdiff(1:n, ind.train)

  # Get results from all methods
  res <- bind_rows(
    PRS(G, CHR, POS, pheno.all, covar.all, ind.train, ind.test),
    logit.CMSA(G,  pheno.all, covar.all, ind.train, ind.test, "logit-simple"),
    logit.CMSA(G2, pheno.all, covar.all, ind.train, ind.test, "logit-triple"),
    ttrees(TTree, "backingfiles/ttrees", pheno.all, ind.train, ind.test, n.trees = 100)
  )
  params[["res"]] <- list(res)
  saveRDS(unnest(params, res, .drop = FALSE), file = res.file)
}
```

### Scenario n°1 (without T-Trees)

```{r}
params.grid2 <- expand.grid(
  n.train = 6000,
  par.causal = list(c(30, "all"), c(300, "all"), c(3000, "all"), c(30, "HLA")), 
  par.dist   = c("gaussian", "laplace"), 
  par.h2     = c(0.5, 0.8), 
  par.model  = c("simple", "fancy"),
  num.simu   = 1:100,
  stringsAsFactors = FALSE
) 
```

```{r, eval=TRUE}
if (!dir.exists("results2")) dir.create("results2")

for (i in rows_along(params.grid2)) {

  res.file <- paste0("results2/simu_", i, ".rds")
  if (file.exists(res.file)) next

  params <- params.grid2[i, ]
  par.causal <- params[["par.causal"]][[1]]

  # Simulate phenotypes
  pheno.all <- get_pheno(
    G,    
    h2 = params[["par.h2"]], 
    M = as.integer(par.causal[1]), 
    ind.possible = `if`(par.causal[2] == "all", cols_along(G), ind.HLA),
    effects.dist = params[["par.dist"]], 
    model = params[["par.model"]], 
    K = 0.3                 
  )

  # Split in training/test sets
  ind.train <- sort(sample(n, size = params[["n.train"]]))
  ind.test <- setdiff(1:n, ind.train)

  # Get results from all methods
  res <- bind_rows(
    PRS(G, CHR, POS, pheno.all, covar.all, ind.train, ind.test),
    logit.CMSA(G,  pheno.all, covar.all, ind.train, ind.test, "logit-simple"),
    logit.CMSA(G2, pheno.all, covar.all, ind.train, ind.test, "logit-triple")
  )
  params[["res"]] <- list(res)
  saveRDS(unnest(params, res, .drop = FALSE), file = res.file)
}
```

### Scenario n°2 (dataset with only chromosome 6)

```{r}
params.grid4 <- expand.grid(
  n.train    = 6000,
  par.causal = list(c(30, "all"), c(300, "all"), c(3000, "all"), c(30, "HLA")), 
  par.dist   = c("gaussian", "laplace"), 
  par.h2     = c(0.5, 0.8), 
  par.model  = "simple",
  num.simu   = 1:100,
  stringsAsFactors = FALSE
) 
```

```{r, eval=TRUE}
if (!dir.exists("results4")) dir.create("results4")

G6 <- big_copy(G, ind.col = which(CHR == 6))
ind.HLA6 <- snp_indLRLDR(CHR[CHR == 6], POS[CHR == 6], 
                         subset(LD.wiki34, ID == "hild12"))

for (i in rows_along(params.grid4)) {

  res.file <- paste0("results4/simu_", i, ".rds")
  if (file.exists(res.file)) next

  params <- params.grid4[i, ]
  par.causal <- params[["par.causal"]][[1]]

  # Simulate phenotypes
  pheno.all <- get_pheno(
    G6,    
    h2 = params[["par.h2"]], 
    M = as.integer(par.causal[1]), 
    ind.possible = `if`(par.causal[2] == "all", cols_along(G6), ind.HLA6),
    effects.dist = params[["par.dist"]], 
    model = params[["par.model"]], 
    K = 0.3                 
  )

  # Split in training/test sets
  ind.train <- sort(sample(n, size = params[["n.train"]]))
  ind.test <- setdiff(1:n, ind.train)

  # Get results from all methods
  res <- bind_rows(
    PRS(G6, CHR[CHR == 6], POS[CHR == 6], pheno.all, covar.all, ind.train, ind.test),
    logit.CMSA(G6, pheno.all, covar.all, ind.train, ind.test, "logit-simple")
  )
  params[["res"]] <- list(res)
  saveRDS(unnest(params, res, .drop = FALSE), file = res.file)
}
```

### Scenario n°3 (varying training size)

```{r}
params.grid5 <- expand.grid(
  n.train    = 1:5 * 1000,
  par.causal = list(c(300, "all")), 
  par.dist   = c("gaussian", "laplace"), 
  par.h2     = c(0.5, 0.8), 
  par.model  = "simple",
  num.simu   = 1:100,
  stringsAsFactors = FALSE
) 
```

```{r, eval=TRUE}
if (!dir.exists("results5")) dir.create("results5")

for (i in rows_along(params.grid5)) {

  res.file <- paste0("results5/simu_", i, ".rds")
  if (file.exists(res.file)) next

  params <- params.grid5[i, ]
  par.causal <- params[["par.causal"]][[1]]

  # Simulate phenotypes
  pheno.all <- get_pheno(
    G,    
    h2 = params[["par.h2"]], 
    M = as.integer(par.causal[1]), 
    ind.possible = `if`(par.causal[2] == "all", cols_along(G), ind.HLA),
    effects.dist = params[["par.dist"]], 
    model = params[["par.model"]], 
    K = 0.3                 
  )

  # Split in training/test sets
  ind.train <- sort(sample(n, size = params[["n.train"]]))
  ind.test <- setdiff(1:n, ind.train)

  # Get results from all methods
  res <- bind_rows(
    PRS(G, CHR, POS, pheno.all, covar.all, ind.train, ind.test),
    logit.CMSA(G,  pheno.all, covar.all, ind.train, ind.test, "logit-simple")
  )
  params[["res"]] <- list(res)
  saveRDS(unnest(params, res, .drop = FALSE), file = res.file)
}
```

### Scenario n°1 (comparison with biglasso)

```{r}
library(biglasso)
library(Matrix)
```

```{r, eval=FALSE}
G3 <- big.matrix(nrow(G), ncol(G) + ncol(covar.all),
                 backingfile = "G-PC", 
                 backingpath = "backingfiles")
big_apply(G, function(X, ind) {
  G3[, ind] <- X[, ind]
  NULL
}, a.combine = 'c')
G3[, ncol(G) + cols_along(covar.all)] <- covar.all
```

```{r}
G3 <- attach.big.matrix("backingfiles/G-PC.desc")
```

```{r}
params.grid7 <- expand.grid(
  n.train = 6000,
  par.causal = list(c(30, "all"), c(300, "all"), c(3000, "all"), c(30, "HLA")), 
  par.dist   = c("gaussian", "laplace"), 
  par.h2     = c(0.8), 
  par.model  = c("simple"),
  num.simu   = 1:100,
  stringsAsFactors = FALSE
) 
```

```{r, eval=TRUE}
if (!dir.exists("results7")) dir.create("results7")

for (i in rows_along(params.grid7)) {

  res.file <- paste0("results7/simu_", i, ".rds")
  if (file.exists(res.file)) next

  params <- params.grid7[i, ]
  par.causal <- params[["par.causal"]][[1]]

  # Simulate phenotypes
  pheno.all <- get_pheno(
    G,    
    h2 = params[["par.h2"]], 
    M = as.integer(par.causal[1]), 
    ind.possible = `if`(par.causal[2] == "all", cols_along(G), ind.HLA),
    effects.dist = params[["par.dist"]], 
    model = params[["par.model"]], 
    K = 0.3                 
  )

  # Split in training/test sets
  ind.train <- sort(sample(n, size = params[["n.train"]]))
  ind.test <- setdiff(1:n, ind.train)

  # Get results from all methods
  res <- bind_rows(
    logit.CMSA(G, pheno.all, covar.all, ind.train, ind.test, "logit-simple"),
    logit.biglasso(G3, pheno.all, covar.all, ind.train, ind.test)
  )
  params[["res"]] <- list(res)
  saveRDS(unnest(params, res, .drop = FALSE), file = res.file)
}
```