### 2018-08-01 4-results-simus.Rmd
analyze new results

---
title: "Results"
author: "Florian Privé"
date: "November 17, 2017"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", out.width = "70%", fig.width = 10,
                      fig.asp = 0.75, dev = "svg")
options(width = 110)
```

```{r}
library(tidyverse)
# devtools::install_github("privefl/bigsnpr")
library(bigsnpr)
# devtools::install_github("privefl/paper2-PRS/pkg.paper.PRS")
library(pkg.paper.PRS)  ## See code source for functions' code.
options(nboot = 1e4)
```

## Useful functions

```{r, message=FALSE}
COLORS <- scales::hue_pal()(6)
methods.color <- setNames(
  c(COLORS, COLORS[c(4, 2, 5)]), 
  c("logit-simple", "PRS-max", "PRS-stringent", "T-Trees", "logit-triple", "PRS-all",
    "PRS-max-0.05", "PRS-max-0.2", "PRS-max-0.8")
) 
```

```{r}
myggplot <- function(..., coeff = 1) {
  ggplot(...) + 
    bigstatsr::theme_bigstatsr(size.rel = coeff) +
    theme(strip.text.x = element_text(size = rel(2)),
          strip.text.y = element_text(size = rel(2)))
} 
```

```{r}
barplot_causal <- function(results) {

  results %>%
    myggplot(aes(par.causal, AUC_mean, fill = method, color = method)) +
    geom_hline(yintercept = 0.5, linetype = 2) +
    geom_col(position = position_dodge(), alpha = 0.3) +
    geom_errorbar(aes(ymin = AUC_mean - 2 * AUC_boot, ymax = AUC_mean + 2 * AUC_boot),
                  position = position_dodge(width = 0.9), color = "black", width = 0.2) + 
    scale_y_continuous(limits = c(0.5, NA), minor_breaks = 0:20 / 20,
                       oob = scales::rescale_none) +
    labs(x = "Causal SNPs (number and location)", y = "Mean of 100 AUCs",
         fill = "Method", color = "Method") +
    scale_fill_manual(values = methods.color) +
    scale_color_manual(values = methods.color)
}

barplot_causal_one <- function(results, h2) {

  auc_max <- `if`(h2 == 0.8, 0.94, 0.84)

  results %>%
    filter(par.h2 == h2) %>%
    barplot_causal() +
    geom_hline(yintercept = auc_max, linetype = 3, color = "blue") +
    facet_grid(par.dist ~ .)
}

barplot_causal_all <- function(results) {

  results %>%
    barplot_causal() +
    geom_hline(aes(yintercept = auc_max), linetype = 3, color = "blue",
               data = data.frame(par.h2 = c(0.5, 0.8), auc_max = c(0.84, 0.94))) +
    facet_grid(par.dist ~ par.h2)
}
```

```{r}
compare_logit <- function(results, with_method) {

  results <- filter(
    results, method %in% c("logit-simple", with_method), par.h2 == 0.8)

  results.summary <- results %>%
    group_by_at(c(vars(starts_with("par")), "method")) %>%
    summarise(AUC_mean = mean(AUC), AUC_boot = boot(AUC), N = n())

  nsimu <- results.summary$N[[1]]
  # stopifnot(all(results.summary$N == nsimu))

  plot_extra <- function(y, ylab = y) {

    results %>%
      myggplot() +
      geom_boxplot(aes_string("par.causal", y, color = "method", 
                              fill = "method"), alpha = 0.3) + 
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      facet_grid(par.model ~ par.dist) +
      labs(x = "Causal SNPs (number and location)", y = ylab, 
           fill = "Method", color = "Method") +
      theme(legend.position = "none") +
      scale_fill_manual(values = methods.color) +
      scale_color_manual(values = methods.color) +
      scale_y_continuous(limits = c(0, NA))
  }

  p1 <- results.summary %>%
    barplot_causal_one(h2 = 0.8) +
    facet_grid(par.model ~ par.dist) +
    labs(y = sprintf("Mean of %d AUCs", nsimu))

  p2 <- cowplot::plot_grid(
    plot_extra("nb.preds", "Number of predictors"),
    plot_extra("timing", "Execution time (in seconds)"),
    ncol = 1, align = "hv", scale = 0.95, 
    labels = LETTERS[2:3], label_size = 20
  )

  p12 <- cowplot::plot_grid(
    p1 + theme(legend.position = "none"), p2, 
    ncol = 2, scale = 0.95, rel_widths = c(6, 4), 
    labels = c("A", ""), label_size = 20
  )

  cowplot::plot_grid(
    cowplot::get_legend(p1 + theme(legend.direction = "horizontal")), p12, 
    rel_heights = c(0.1, 1), ncol = 1
  )
}
```

```{r, eval=FALSE}
G <- snp_attach("backingfiles/celiacQC_sub1.rds")$genotypes
corr <- snp_cor(G, size = 100)
saveRDS(corr, "backingfiles/corr2.rds")
```

```{r}
corr <- readRDS("backingfiles/corr2.rds")
```

## Results with T-Trees

```{r}
results1 <- list.files("results1", full.names = TRUE) %>%
  read_format_results(corr)
```

```{r, out.width="90%", fig.width = 14}
compare_logit(results1, with_method = "T-Trees")

ggsave("figures/supp-ttrees.pdf", scale = 1/100, width = 1400, height = 1060)
```

## Results with logit-triple & model fancy

```{r}
results2 <- list.files("results2", full.names = TRUE) %>%
  read_format_results(corr)
```

### Results with logit-triple

```{r, out.width="90%", fig.width = 14}
compare_logit(results2, with_method = "logit-triple")

ggsave("figures/supp-triple.pdf", scale = 1/100, width = 1400, height = 1060)
```

```{r}
results2 %>%
  filter(grepl("logit", method)) %>%
  group_by_at(c(vars(starts_with("par")), "method")) %>%
  summarise(AUC_mean = mean(AUC), AUC_boot = boot(AUC), N = n()) %>%
  barplot_causal_one(h2 = 0.5) +
  facet_grid(par.model ~ par.dist)

ggsave("figures/supp-AUC-triple.pdf", scale = 1/110, width = 1242, height = 951)
```

### Results with model fancy

```{r}
results2 %>%
  filter(par.model == "fancy", method %in% c("logit-simple", "PRS-max")) %>%
  filter(thr.r2 == 0.2 | is.na(thr.r2)) %>%
  group_by_at(c(vars(starts_with("par")), "method")) %>%
  summarise(AUC_mean = mean(AUC), AUC_boot = boot(AUC), N = n()) %>%
  barplot_causal_all()

ggsave("figures/supp-AUC-logit-fancy.pdf", scale = 1/110, width = 1242, height = 951)
```

```{r}
results2 %>%
  filter(par.model == "fancy", grepl("PRS", method)) %>%
  filter(thr.r2 == 0.2 | is.na(thr.r2)) %>%
  group_by_at(c(vars(starts_with("par")), "method")) %>%
  summarise(AUC_mean = mean(AUC), AUC_boot = boot(AUC), N = n()) %>%
  barplot_causal_all()

ggsave("figures/supp-AUC-PRS-fancy.pdf", scale = 1/110, width = 1242, height = 951)
```

## Results with only logit-simple & PRSs

### Different r2 thresholds

```{r}
results2 %>%
  filter(par.model == "simple", method %in% c("logit-simple", "PRS-max")) %>%
  mutate(method = c_method_r2(method, thr.r2)) %>%
  group_by_at(c(vars(starts_with("par")), "method")) %>%
  summarise(AUC_mean = mean(AUC), AUC_boot = boot(AUC), N = n()) %>%
  barplot_causal_all()

ggsave("figures/main-AUC-all-r2.pdf", scale = 1/110, width = 1242, height = 951)
```

```{r}
results2 %>%
  filter(par.model == "simple", method %in% c("logit-simple", "PRS-max")) %>%
  mutate(method = c_method_r2(method, thr.r2)) %>%
  group_by_at(c(vars(starts_with("par")), "method")) %>%
  summarise(AltSens_mean = mean(AltSens), AltSens_boot = boot(AltSens), N = n()) %>%
  myggplot(aes(par.causal, AltSens_mean, fill = method, color = method)) + 
  geom_col(position = position_dodge(), alpha = 0.3) +
  geom_errorbar(aes(ymin = AltSens_mean - 2 * AltSens_boot,
                    ymax = AltSens_mean + 2 * AltSens_boot),
                position = position_dodge(width = 0.9), color = "black", width = 0.2) + 
  scale_y_continuous(limits = c(0, 1)) +
  labs(x = "Causal SNPs (number and location)", y = "Mean of 100 AltSens",
       fill = "Method", color = "Method") +
  scale_fill_manual(values = methods.color) +
  scale_color_manual(values = methods.color) +
  facet_grid(par.dist ~ par.h2)

ggsave("figures/supp-AltSens-all-r2.pdf", scale = 1/110, width = 1242, height = 951)
```

### Correlation between predictive measures

```{r}
results3 <- results2 %>%
  filter(thr.r2 == 0.2 | is.na(thr.r2)) %>%
  filter(par.model == "simple", method != "logit-triple")
```

```{r, fig.asp=0.7, out.width="95%", fig.width=16}
results3 %>%
  filter(method %in% c("logit-simple", "PRS-max")) %>%
  mutate(Parameters  = interaction(par.h2, par.dist, method, sep = " | ")) %>%
  myggplot(aes(AUC, pAUC, color = Parameters)) + 
  geom_point(size = 0.6, alpha = 0.5) +
  geom_smooth(aes(linetype = Parameters), method = "loess", se = FALSE, size = 2) +
  theme(legend.key.width = unit(3.5, "line"), 
        legend.direction = "horizontal") + 
  labs(y = "Partial AUC (specificity: [0.9, 1])") +
  theme(legend.position = "top", axis.title = element_text(size = rel(2.2))) +

ggsave("figures/supp-AUC-corr.pdf", scale = 1/100, width = 1525, height = 983)
```

```{r}
cor(results3$AUC, results3$pAUC, method = "spearman")
```

### 6000 in training & all chromosomes

```{r}
results3.summary <- results3 %>%
  group_by_at(c(vars(starts_with("par")), "method")) %>%
  summarise(AUC_mean = mean(AUC), AUC_boot = boot(AUC), N = n())
```

```{r, fig.asp = 1, out.width = "60%", fig.width=8}
results3.summary %>%
  filter(method %in% c("logit-simple", "PRS-max"), par.model == "simple") %>%
  barplot_causal_one(h2 = 0.8)

ggsave("figures/main-AUC-logit.pdf", scale = 1/90, width = 844, height = 872)
```

```{r, fig.asp = 1, out.width = "60%", fig.width=8}
results3.summary %>%
  filter(grepl("PRS", method)) %>%
  barplot_causal_one(h2 = 0.8)

ggsave("figures/main-AUC-PRS.pdf", scale = 1/90, width = 844, height = 872)
```

```{r, fig.asp = 1, out.width = "60%", fig.width=8}
results3.summary %>%
  filter(method %in% c("logit-simple", "PRS-max")) %>%
  barplot_causal_one(h2 = 0.5)

ggsave("figures/supp-AUC-logit.pdf", scale = 1/90, width = 844, height = 872)
```

```{r, fig.asp = 1, out.width = "60%", fig.width=8}
results3.summary %>%
  filter(grepl("PRS", method)) %>%
  barplot_causal_one(h2 = 0.5)

ggsave("figures/supp-AUC-PRS.pdf", scale = 1/90, width = 844, height = 872)
```

### 6000 in training & only chromosome 6

```{r}
results4 <- list.files("results4", full.names = TRUE) %>%
  read_format_results(corr)
```

```{r, fig.asp = 1, out.width = "90%"}
results4 %>%
  filter(method %in% c("logit-simple", "PRS-max")) %>%
  mutate(method = c_method_r2(method, thr.r2)) %>%
  group_by_at(c(vars(starts_with("par")), "method")) %>%
  summarise(AUC_mean = mean(AUC), AUC_boot = boot(AUC), N = n()) %>%
  barplot_causal_all()

ggsave("figures/supp-AUC-chr6-all-r2.pdf", scale = 1/110, width = 1242, height = 951)
```

### Varying training size & all shromosomes

```{r}
results5 <- list.files("results5", full.names = TRUE) %>%
  read_format_results(corr)

results6 <- results2 %>%
  filter(par.model == "simple", par.causal == "300 in all") %>%
  bind_rows(results5) %>%
  filter(method %in% c("logit-simple", "PRS-max")) %>%
  mutate(method = c_method_r2(method, thr.r2))
```

```{r}
results6 %>%
  group_by_at(c(vars(starts_with("par")), "method", "n.train")) %>%
  summarise(AUC_mean = mean(AUC), AUC_boot = boot(AUC), N = n()) %>%
  myggplot(aes(n.train, AUC_mean, color = method, linetype = method)) +
  geom_hline(yintercept = 0.5, linetype = 2) +
  geom_hline(aes(yintercept = auc_max), linetype = 3, color = "blue",
             data = data.frame(par.h2 = c(0.5, 0.8), auc_max = c(0.84, 0.94))) +
  geom_line(size = 1.5) +
  geom_errorbar(aes(ymin = AUC_mean - 2 * AUC_boot, 
                    ymax = AUC_mean + 2 * AUC_boot), 
                size = 1.5, width = 0) +
  facet_grid(par.dist ~ par.h2) +
  scale_x_continuous(breaks = 1:6 * 1000, minor_breaks = NULL) + 
  scale_y_continuous(minor_breaks = 0:20 / 20) +
  scale_color_manual(values = methods.color) +
  labs(x = "Size of the training set", y = "Mean of 100 AUCs", 
       color = "Method", linetype = "Method") +
  theme(legend.key.width = unit(3, "line"))

ggsave("figures/main-AUC-ntrain.pdf", scale = 1/100, width = 1050, height = 800)
```