### 2018-09-27 results-simus-pres.Rmd
new figures for pres

---
title: "Results"
author: "Florian Privé"
date: "September 27, 2018"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", out.width = "70%", fig.width = 10,
                      fig.asp = 0.75, dev = "svg")
options(width = 110)
```

## Useful functions

```{r, message=FALSE}
library(tidyverse)
library(pkg.paper.PRS)
options(nboot = 1e4)

COLORS <- scales::hue_pal()(6)
methods.color <- setNames(
  c(COLORS, COLORS[c(4, 2, 5)], "black"), 
  c("PLR", "C+T-max", "C+T-stringent", "T-Trees", "PLR3", "C+T-all",
    "C+T-max-0.05", "C+T-max-0.2", "C+T-max-0.8", "biglasso")
) 
```

```{r}
myggplot <- function(..., coeff = 1) {
  ggplot(...) + bigstatsr::theme_bigstatsr(size.rel = coeff)
} 
```

```{r}
barplot_causal <- function(results) {

  results %>%
    myggplot(aes(par.causal, AUC_mean, fill = method, color = method)) +
    geom_hline(yintercept = 0.5, linetype = 2) +
    geom_col(position = position_dodge(), alpha = 0.5, color = "black", size = 1) +
    geom_errorbar(aes(ymin = AUC_mean - 2 * AUC_boot, ymax = AUC_mean + 2 * AUC_boot),
                  position = position_dodge(width = 0.9), color = "black", width = 0.2, size = 1) + 
    scale_y_continuous(limits = c(0.5, NA), minor_breaks = 0:20 / 20,
                       oob = scales::rescale_none) +
    labs(x = "Causal SNPs (number and location)", y = "Mean of 100 AUCs",
         fill = "Method", color = "Method") +
    scale_fill_manual(values = methods.color) +
    scale_color_manual(values = methods.color)
}

barplot_causal_one <- function(results) {

  results %>%
    barplot_causal() +
    geom_hline(yintercept = 0.94, linetype = 3, color = "blue")
}
```

```{r}
compare_PLR <- function(results, with_method) {

  results <- filter(results, method %in% c("PLR", with_method))

  results.summary <- results %>%
    group_by_at(c(vars(starts_with("par")), "method")) %>%
    summarise(AUC_mean = mean(AUC), AUC_boot = boot(AUC), N = n())

  nsimu <- results.summary$N[[1]]
  # stopifnot(all(results.summary$N == nsimu))

  plot_extra <- function(y, ylab = y) {

    results %>%
      myggplot() +
      geom_boxplot(aes_string("par.causal", y, color = "method", 
                              fill = "method"), alpha = 0.3) + 
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      facet_grid(par.model ~ .) +
      labs(x = "Causal SNPs (number and location)", y = ylab, 
           fill = "Method", color = "Method") +
      theme(legend.position = "none") +
      scale_fill_manual(values = methods.color) +
      scale_color_manual(values = methods.color) +
      scale_y_continuous(limits = c(0, NA))
  }

  p1 <- results.summary %>%
    barplot_causal_one() +
    facet_grid(par.model ~ .) +
    labs(y = sprintf("Mean of %d AUCs", nsimu))

  p2 <- cowplot::plot_grid(
    plot_extra("nb.preds", "Number of predictors"),
    plot_extra("timing", "Execution time (in seconds)"),
    ncol = 1, align = "hv", scale = 0.95, 
    labels = LETTERS[2:3], label_size = 20
  )

  p12 <- cowplot::plot_grid(
    p1 + theme(legend.position = "none"), p2, 
    ncol = 2, scale = 0.95, rel_widths = c(6, 4), 
    labels = c("A", ""), label_size = 20
  )

  cowplot::plot_grid(
    cowplot::get_legend(p1 + theme(legend.direction = "horizontal")), p12, 
    rel_heights = c(0.1, 1), ncol = 1
  )
}
```

```{r}
corr <- readRDS("backingfiles/corr2.rds")
```

## Results with logit-triple & model fancy

```{r}
results2 <- list.files("results2", full.names = TRUE) %>%
  read_format_results(corr) %>%
  filter(par.h2 == 0.8, par.dist == "Gaussian")
```

### Results with PLR3

```{r, out.width="90%", fig.width = 14}
compare_PLR(results2, with_method = "PLR3")

ggsave("../thesis-docs/figures/pres2-triple.svg", scale = 1/100, width = 1300, height = 1060)
```

## Results with only logit-simple & PRSs

### 6000 in training & all chromosomes

```{r}
results2.summary <- results2 %>%
  filter(par.model == "ADD", method != "PLR3") %>%
  group_by_at(c(vars(starts_with("par")), "method")) %>%
  summarise(AUC_mean = mean(AUC), AUC_boot = boot(AUC), N = n())
```

```{r}
results2.summary %>%
  filter(method %in% c("PLR", "C+T-max")) %>%
  barplot_causal_one()

ggsave("../thesis-docs/figures/pres2-AUC-logit.svg", scale = 1/110, width = 1045, height = 770)
```

```{r}
results2.summary %>%
  filter(grepl("C\\+T", method)) %>%
  barplot_causal_one()

ggsave("../thesis-docs/figures/pres2-AUC-PRS.svg", scale = 1/110, width = 1045, height = 770)
```

### 6000 in training & only chromosome 6

```{r}
results4 <- list.files("results4", full.names = TRUE) %>%
  read_format_results(corr) %>%
  filter(par.h2 == 0.8, par.dist == "Gaussian")

results4.summary <- results4 %>%
  filter(method %in% c("PLR", "C+T-max")) %>%
  mutate(method = c_method_r2(method, thr.r2)) %>%
  group_by_at(c(vars(starts_with("par")), "method")) %>%
  summarise(AUC_mean = mean(AUC), AUC_boot = boot(AUC), N = n())

results2.all.r2 <- results2 %>%
  filter(par.model == "ADD", method %in% c("PLR", "C+T-max")) %>%
  mutate(method = c_method_r2(method, thr.r2)) %>%
  group_by_at(c(vars(starts_with("par")), "method")) %>%
  summarise(AUC_mean = mean(AUC), AUC_boot = boot(AUC), N = n())
```

```{r}
results4.summary %>%
  barplot_causal_one() + 
  geom_col(aes(par.causal, AUC_mean), data = results2.all.r2, 
           position = position_dodge(), color = "black", alpha = 0)

ggsave("../thesis-docs/figures/pres2-AUC-chr6.svg", scale = 1/110, width = 990, height = 700)
```

### Varying training size & all shromosomes

```{r}
results5 <- list.files("results5", full.names = TRUE) %>%
  read_format_results(corr) %>%
  filter(par.h2 == 0.8, par.dist == "Gaussian")

results5.summary <- results5 %>%
  group_by_at(c(vars(starts_with("par")), "method", "n.train")) %>%
  summarise(AUC_mean = mean(AUC), AUC_boot = boot(AUC), N = n())

results35.summary <- results2.summary %>%
  cbind.data.frame(n.train = 6000) %>%
  bind_rows(results5.summary) %>%
  filter(par.causal == "300 in all")
```

```{r}
results35.summary %>%
  filter(method %in% c("PLR", "C+T-max")) %>%
  myggplot(aes(n.train, AUC_mean, color = method, linetype = method)) +
  geom_hline(yintercept = 0.5, linetype = 2) +
  geom_hline(aes(yintercept = 0.94), linetype = 3, color = "blue") +
  geom_line(size = 1.5) +
  geom_errorbar(aes(ymin = AUC_mean - 2 * AUC_boot, 
                    ymax = AUC_mean + 2 * AUC_boot), 
                size = 1.5, width = 0) +
  scale_x_continuous(breaks = 1:6 * 1000, minor_breaks = NULL) + 
  scale_y_continuous(minor_breaks = 0:20 / 20) +
  scale_color_manual(values = methods.color) +
  labs(x = "Size of the training set", y = "Mean of 100 AUCs", 
       color = "Method", linetype = "Method") +
  theme(legend.key.width = unit(3, "line"))

ggsave("../thesis-docs/figures/pres2-AUC-ntrain.svg", scale = 1/100, width = 870, height = 635)
```